Can Zero-shot LLMs save additional work in machine learning prioritised screening for systematic reviews?

### 2024-11-07 LLMs/manuscript/main.tex
Updates from Overleaf

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[margin=2.5cm]{geometry}

@@

	\begin{abstract}
		Zero shot learning using LLMs has received widespread attention in recent years, including in the automation of systematic review screening. 
		However, there has been insufficient research on robust processes to automate screening responsibly. 
		For systematic review screening, responsible automation requires that the risks of missing relevant studies (which are inherent when the process is automated) can be accurately quantified and managed.
		In this paper, we adapt LLM-assisted screening to the widespread machine learning prioritised screening paradigm, and use robust stopping criteria to quantify how much work can be responsibly saved with different levels of risk tolerance for missing studies.
		We find that - even in the absence of detailed inclusion and exclusion criteria - LLM-assisted screening can produce ranked lists of documents that are superior to those generated by the baseline.
		However, LLMs do not seem to be a silver bullet which would substantially reduce the amount of work required to complete screening. Across a variety of recall targets and confidence levels, work savings were generally higher with the SVM baseline, though possible work savings with LLMS were higher in low recall or low confidence settings. 
	\end{abstract}

@@

	Systematic Reviews are important tools to synthesise findings across multiple studies, enabling decision-makers to make decisions that are informed by the best available evidence. 
	A key part of the process of systematic reviews involves finding studies using a transparent process that aims to be as comprehensive as possible. 
	Usually, bibliographic records are searched for using structured queries across multiple databases \cite{lefebvre_chapter_2023}.
	This results in a list of records that may or not be relevant to the review. The titles and abstracts of these records are then examined by hand - often by two independent reviewers - and it is judged whether they should be included, which means they are read again at the full text level, or excluded.
	This is a laborious and repetitive process - known as screening - and in times of increasing numbers of publications, this becomes more and more challenging.

	There is a large literature \cite{omara-eves_using_2015}, going back nearly 20 years \cite{cohen_reducing_2006}, that investigates how machine learning might be used to automate parts of this process, in order to reduce the labour required to complete a systematic review. However, much of this literature introduces new tools, or demonstrates how one approach or another \textit{would have been able to} save work on past reviews where complete evaluation datasets are available, without showing how the approach could be used to \textit{responsibly} save work in new reviews where the inclusion and exclusion decisions are not known \textit{a priori}  \cite{van_de_schoot_open_2021, rathbone_faster_2015, yang_goldilocks_2022}.

	This is reflected in the fact that the guidelines for conducting systematic reviews published by the Cochrane colloquium \cite{lefebvre_chapter_2023} do not recommend that machine learning is used to save labour during screening. This is because it remains uncertain how the use of automation may result in relevant studies being missed.

	Early attempts to automate systematic review screening used simple models like logistic regression and support vector machines \cite{cohen_reducing_2006}. Lately, the literature on automating systematic reviews has increasingly turned its attention to the use of large language models (LLMs) \cite{oami_performance_2024, wang_zero-shot_2024, luo_evaluating_2024, xia_llmscreen_2024}. However, not enough attention has been paid to comparing LLM-based approaches with the much simpler (and computationally less expensive) alternatives. Further, papers which have investigated the use of LLMs have so far not considered how they might be used in a responsible manner where evaluation data is not available.

@@

	Although the LLM-based procedure is compared to a fine-tuned BERT-like model, the details of this are not clear, and it appears that there is no comparison to the most common mode of machine learning assisted screening, known as machine learning prioritised screening. With ML-prioritised screening, a subset of documents are screened by hand, and these documents are used to train a model to predict the relevance of remaining unscreened documents. A batch of those records with the highest predictions are then screened by hand, before the model is retrained using the expanded training dataset. This process continues until a ``stopping criterion'' \cite{SneydS19, callaghan_statistical_2020, lewis_confidence_2023} is met, which either aims to guarantee that a certain recall level has been met with a given confidence level, or which quantifies that the returns to additional screening are low.

@@

	In this paper, we adapt the probabilistic zero-shot screening process using LLMs to a ML-prioritised screening setting, and show that - when combined with statistically valid stopping criterion - it is able to reliably achieve high levels of recall while saving work. Secondly, we compare LLM-assisted approaches to simpler methods available, both in terms of the work saved, and in terms of computational requirements.

@@

	In a new review, one could generate scores in this way for each document returned by a query, and screen these in descending order of their inclusion score, applying a stopping criterion to decide when remaining unscreened documents can be excluded automatically.  We simulate this process using the Synergy dataset of completed reviews \cite{de_bruin_synergy_2023}, and report on the work that would have been saved, and the actual level of recall that would have been achieved, given a range of recall targets and uncertainty levels.

@@

	We implement our active learning pipeline as follows for each model and each review dataset, where each dataset has $N$ unique documents. Screening here, in a simulation context, refers to revealing the screening decision that was made when the dataset was created.

@@

	Overall, across the X datasets, Llama3.1 models achieves a higher ROC AUC score than the SVM baseline, while Llama2 models achieve a lower score (figure \ref{fig:macro}). Across both generations of Llama models, models with more parameters achieved higher scores than models with fewer paramaters. This difference was more pronounced for Llama2 than for Llama3.1.

@@

	These results add to previous evidence that it is possible to \textit{safely} use prioritised screening approaches that can guarantee a given level of recall with a given level of confidence \cite{callaghan_statistical_2020}. Across a wide range of datasets, as well as a wide range of different models, the stopping criteria ensured that the targeted level of recall was missed no more than expected given the confidence level.

@@

	Simplified version, no inclusion criteria - but also simplified SVM setting, no parameter tuning, large batches, no finetuned.

	Possibility to combine

	Scientific treatment of the costs and benefits of different options.

	\bibliographystyle{unsrt}

	\bibliography{LLMs/stopping-criteria}
	%\bibliography{../LLMs}

>>>>

\documentclass{article}
\usepackage[utf8]{inputenc}
\usepackage{authblk}
\usepackage{setspace}
\usepackage[maxbibnames=5]{biblatex}
%\usepackage{natbib}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage[margin=2.5cm]{geometry}

\addbibresource{LLMs/stopping-criteria.bib}

@@

	\begin{abstract}
		Zero shot learning using LLMs has received widespread attention in recent years, including in the automation of systematic review screening. 
		However, there has been insufficient research on robust processes to automate screening responsibly. 
		For systematic review screening, responsible automation requires that the risks of missing relevant studies (which are inherent when the process is automated) can be accurately quantified and managed. 
		In this paper, we adapt LLM-assisted screening to the widespread machine learning prioritised screening paradigm, and use a robust stopping criterion to quantify how much work can be responsibly saved with different levels of risk tolerance for missing studies.
		We find that - even in the absence of detailed inclusion and exclusion criteria - LLM-assisted screening can produce ranked lists of documents that are superior to those generated by our baseline - a standard active learning approach using a support vector machine.
		However, LLMs do not seem to be a silver bullet which would substantially reduce the amount of work required to complete screening. Across a variety of recall targets and confidence levels, work savings were generally higher with the SVM baseline, though possible work savings with LLMs were higher in low recall or low confidence settings. 
	\end{abstract}

@@

	Rigorous systematic reviews methodologies are critical to synthesise findings across multiple studies, enabling decisions to be informed by the best available evidence. 
	A key part of the systematic review process involves finding studies using a transparent process that aims to be as comprehensive as possible. 
	Usually, bibliographic records are searched for using structured queries across multiple databases \cite{lefebvre_chapter_2023}.
	This results in a list of records that may or not be relevant to the review. The titles and abstracts of these records are then examined by hand - often by two independent reviewers - and it is judged whether they should be included, which means they are read again at the full text level, or excluded, based on a transparent set of criteria demarcating the study boundaries.
	This is a laborious and repetitive process - known as screening - and in times of increasing numbers of publications, this becomes more and more challenging \cite{bornmann_growth_2015}.

	There is a large literature \cite{omara-eves_using_2015}, going back nearly 20 years \cite{cohen_reducing_2006}, that investigates how machine learning might be used to automate parts of the screening process, in order to reduce the labour required to complete a systematic review. However, much of this literature introduces new tools, or demonstrates how one approach or another \textit{would have been able to} save work on past reviews where complete evaluation datasets are available, without showing how the approach could be used to \textit{responsibly} save work in new reviews where the inclusion and exclusion decisions are not known \textit{a priori}  \cite{van_de_schoot_open_2021, rathbone_faster_2015, yang_goldilocks_2022}. This is insufficient. Evaluations in realistic settings are necessary for digital tools for the automation of systematic reviews to be used responsibly.

	This is reflected in the fact that the guidelines for conducting systematic reviews published by the Cochrane collaboration \cite{lefebvre_chapter_2023} do not recommend that machine learning is used to save labour during screening. This is because it remains uncertain how the use of automation may result in relevant studies being missed. This uncertainty will remain until the community reaches a consensus on what an adequate stopping criterion would look like, and what levels of confidence would be adequate to secure methodological standards.

	Early attempts to accelerate systematic review screening used simple models like logistic regression and support vector machines \cite{cohen_reducing_2006} in an active learning, or prioritised screening, setting. In such a setting, a subset of documents are screened by hand, and these documents are used to train a model to predict the relevance of remaining unscreened documents. A batch of those records with the highest predictions are then screened by hand, before the model is retrained using the expanded training dataset. This process continues until a ``stopping criterion'' \cite{SneydS19, callaghan_statistical_2020, lewis_confidence_2023} is met, which either aims to guarantee that a certain recall level has been met with a given confidence level, or which quantifies that the returns to additional screening are low.

 Lately, the literature on automating systematic reviews has increasingly turned its attention to the use of large language models (LLMs) \cite{oami_performance_2024, wang_zero-shot_2024, luo_evaluating_2024, xia_llmscreen_2024}. However, not enough attention has been paid to comparing LLM-based approaches with the much simpler (and computationally less expensive) alternatives. Further, papers which have investigated the use of LLMs have so far not considered how they might be used in a responsible manner where evaluation data is not available.

@@

	Although the LLM-based procedure is compared to a fine-tuned BERT-like model, the details of this are not clear, and it appears that there is no comparison to the most common mode of machine learning assisted screening, prioritised screening.

@@

	In this paper, we adapt the probabilistic zero-shot screening process using LLMs to a ML-prioritised screening setting, and show that - when combined with statistically valid stopping criterion - it is able to reliably achieve high levels of recall while saving work. Secondly, we compare LLM-assisted approaches to simpler methods available.

@@

	In a new review, one could generate scores in this way for each document returned by a query, and screen these in descending order of their inclusion score, applying a stopping criterion to decide when remaining unscreened documents can be excluded automatically.  We simulate this process using the Synergy dataset of completed reviews \cite{de_bruin_synergy_2023}, and report on the work that would have been saved, and the actual level of recall that would have been achieved, given a range of recall targets and uncertainty levels. We do this using small (7/8 billion parameters) and medium (70 billion parameters) sized models from different generations of the Llama family of models \cite{touvron_llama_2023, dubey_llama_2024}. Llama models are open-weights, meaning they can freely be used and re-used, as opposed to proprietary models like GPT4.

@@

	We implement our active learning pipeline as follows for each model and each review dataset, where each dataset has $N$ unique documents. Screening, in this simulation context, refers to revealing the screening decision that was made when the dataset was created.

@@

	Overall, across the 25 systematic review datasets, Llama3.1 models achieves a higher ROC AUC score than the SVM baseline, while Llama2 models achieve a lower score (figure \ref{fig:macro}). Across both generations of Llama models, models with more parameters achieved higher scores than models with fewer parameters. This difference was more pronounced for Llama2 than for Llama3.1.

@@

	These results add to previous evidence that it is possible to \textit{safely} use prioritised screening approaches that, when combined with a robust stopping criterion, can guarantee a given level of recall with a given level of confidence \cite{callaghan_statistical_2020}. Across a wide range of datasets, as well as a wide range of different models, the stopping criteria ensured that the targeted level of recall was missed no more than expected given the confidence level.

@@

    Considering the environmental impact of ever larger models \cite{luccioni_power_2024}, these results merit a dose of caution before the evidence synthesis community rushes to adopt LLMs where simpler models may perform just as well. However, this is not to say that these approaches do not hold promise.

    The results shown here are the product of a very simple implementation of LLM-assisted screening. That they were so impressive with a single zero-shot prediction for each document, using a very simple prompt that contains no additional information about the inclusion and exclusion criteria, hints that the use of LLMs may one day be able to generate substantial work savings. Future research may investigate the effects of using inclusion and exclusion criteria in the prompt, of more extensive "prompt tuning", and incorporating human labels into the prompt via in-context learning \cite{dong_survey_2024}.

    Ultimately though, the utility of prioritisation improvements achieved by more complex models must be balanced against the financial and environmental costs of their deployment, as compared to simpler versions. It must be said also, that the baseline comparison - which consisted of few training iterations, a simple model, and no hyperparameter tuning - was also simplistic.

    Whatever approach is used, the question of how to mitigate the risks of automation to the quality of the study must be paramount. This is a question of statistics, not of machine learning engineering. Currently, prioritised screening with robust stopping criteria represents the best available method to manage the risk of missing studies. The stopping criteria we use has the benefit of ensuring that, should the model used perform less well than expected (but not worse than chance), it is only work savings that will be impacted, while the desired recall level is secured with the desired confidence level.

	%\bibliographystyle{unsrt}

	%\bibliography{LLMs/stopping-criteria}
	%\bibliography{../LLMs}

        \printbibliography
